Human judgement forecasting of COVID-19 in the UK

### 2024-03-18 Analysis.rmd
small updates

(p_data <- plot_data_simple(weekly_truth, daily_truth))
ggsave("output/figures/overview-data.png", plot = p_data, 
       width = 7, height = 5)

>>>>

(p_data <- plot_data_simple(weekly_truth, daily_truth))
ggsave("output/figures/overview-data.png", plot = p_data, 
       width = 7, height = 5, dpi = 300)
ggsave("output/figures/Figure1.tif", plot = p_data, 
       width = 7, height = 5, dpi = 300)

### 2024-03-18 manuscript/manuscript.tex
small updates

Forecasts were elicited from experts and laypeople as part of a public forecasting tournament, the "UK Crowd Forecasting Challenge" using a web application. All forecasts were submitted to the European COVID-19 Forecast Hub, one of several Forecast Hubs that have been systematically collating forecasts of different COVID-19 forecast targets in the US \citep{cramerEvaluationIndividualEnsemble2021}, Germany and Poland \citep{bracherShorttermForecastingCOVID192021, bracherNationalSubnationalShortterm2022}, and Europe \citep{sherrattPredictivePerformanceMultimodel2022}. This study aims to investigate whether the original findings in \citet{bosseComparingHumanModelbased2022} with respect to forecaster performance replicate in a different country, in a different time period, and with an increased number of participants. In addition, it explores the approach proposed in \citet{bosseComparingHumanModelbased2022} to ask participants for a forecast of the estimated effective reproduction number $R_t$ which is then translated into a forecast of cases and deaths using a simulation model. We describe this approach as human in the loop computational modelling and consider it a formalisation of often practiced manual intervention in computational forecasts.

@@

If not otherwise stated, we present results for two-week-ahead forecasts, following the practice adopted by the COVID-19 Forecast Hubs, which found predictive performance to be poor and unreliable beyond this horizon \cite{cramerEvaluationIndividualEnsemble2021, sherrattPredictivePerformanceMultimodel2022, bracherShorttermForecastingCOVID192021}. We analysed all forecasts stratified by forecast target (cases or deaths), forecast horizon, and forecast approach. We compared the performance of the direct vs. $R_t$ forecasting approach using instances where we had both a direct forecasts and an $R_t$ forecast from the same person.

@@

The study period (forecasts were made between May 24 and August 16, 2021, for targets between May 29 and September 11, 2021) was characterised by an increase in the number of cases and deaths in the United Kingdom. Reported cases in particular rose rapidly compared to pre-study levels, with a peak on July 17, 2021, followed by a trough and another subsequent increase in numbers. Death numbers remained almost constant in the first four weeks of the study period, followed by a steady increase until the end of the study period in September 2021. This increase in the case and death numbers coincides with the rise of the Delta variant in the UK at the beginning of May \citep{bastIncreasedRiskHospitalisation2021, perez-guzmanEpidemiologicalDriversTransmissibility2023} as well as the  European Football Championship \citep{dehningImpactEuro20202023}. An overview of the reported case and death numbers is shown in Figure \ref{fig:overview-data}.

@@

% Study period. 
The study period was marked by an increase in both case and death numbers. Case numbers rose quickly compared to the pre-study period, peaking on July 17, 2021, followed by a trough and a subsequent further increase. Forecasts displayed a pattern where forecasters tended to underpredict while case numbers were rising, and overpredict while case numbers were falling, particularly following a peak. Similar patterns have been observed previously in other short-term forecasts of COVID-19 (see e.g, \citep{sherrattPredictivePerformanceMultimodel2022, bracherShorttermForecastingCOVID192021, bosseComparingHumanModelbased2022}).

>>>>

Forecasts were elicited from experts and laypeople as part of a public forecasting tournament, the "UK Crowd Forecasting Challenge" using a web application. All forecasts were submitted to the European COVID-19 Forecast Hub, one of several Forecast Hubs that have been systematically collating forecasts of different COVID-19 forecast targets in the US \citep{cramerEvaluationIndividualEnsemble2021}, Germany and Poland \citep{bracherPreregisteredShorttermForecasting2021a, bracherNationalSubnationalShortterm2022}, and Europe \citep{sherrattPredictivePerformanceMultimodel2022}. This study aims to investigate whether the original findings in \citet{bosseComparingHumanModelbased2022} with respect to forecaster performance replicate in a different country, in a different time period, and with an increased number of participants. In addition, it explores the approach proposed in \citet{bosseComparingHumanModelbased2022} to ask participants for a forecast of the estimated effective reproduction number $R_t$ which is then translated into a forecast of cases and deaths using a simulation model. We describe this approach as human in the loop computational modelling and consider it a formalisation of often practiced manual intervention in computational forecasts.

@@

If not otherwise stated, we present results for two-week-ahead forecasts, following the practice adopted by the COVID-19 Forecast Hubs, which found predictive performance to be poor and unreliable beyond this horizon \cite{cramerEvaluationIndividualEnsemble2021, sherrattPredictivePerformanceMultimodel2022, bracherPreregisteredShorttermForecasting2021a}. We analysed all forecasts stratified by forecast target (cases or deaths), forecast horizon, and forecast approach. We compared the performance of the direct vs. $R_t$ forecasting approach using instances where we had both a direct forecasts and an $R_t$ forecast from the same person.

@@

The study period (forecasts were made between May 24 and August 16, 2021, for targets between May 29 and September 11, 2021) was characterised by an increase in the number of cases and deaths in the United Kingdom. Reported cases in particular rose rapidly compared to pre-study levels, with a peak on July 17, 2021, followed by a trough and another subsequent increase in numbers. Death numbers remained almost constant in the first four weeks of the study period, followed by a steady increase until the end of the study period in September 2021. This increase in the case and death numbers coincides with the rise of the Delta variant in the UK at the beginning of May \citep{bastIncreasedRiskHospitalisation2021, perez-guzmanEpidemiologicalDriversTransmissibility2023} as well as the  European Football Championship \citep{dehningImpactEuro20202023}.Reported cases were likely influenced by an increased uptake of the NHS COVID-19 app in spring and summer 2021 \citep{kendallEpidemiologicalImpactsNHS2023}. An overview of the reported case and death numbers is shown in Figure \ref{fig:overview-data}.

@@

% Study period. 
The study period was marked by an increase in both case and death numbers. Case numbers rose quickly compared to the pre-study period, peaking on July 17, 2021, followed by a trough and a subsequent further increase. Forecasts displayed a pattern where forecasters tended to underpredict while case numbers were rising, and overpredict while case numbers were falling, particularly following a peak. Similar patterns have been observed previously in other short-term forecasts of COVID-19 (see e.g, \citep{sherrattPredictivePerformanceMultimodel2022, bracherPreregisteredShorttermForecasting2021a, bosseComparingHumanModelbased2022}).
